---
title: "data jama"
author: "Chuheng Hu"
date: "November 22, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Upload the file:
```{r}
E <- read.table("extra-activity.csv", sep = ",", header = TRUE)
S <- read.table("student-data.csv", sep = ",", header = TRUE)

```


```{r}
install.packages("party", "rpart")

library(rpart)
library(party)
library(caret)
library(dplyr)
```
```
I want to detect if the parents' education together with who is the guardian could predict the family support

```{r}
#first I select the relevent columns to form a new dataset
FE<-dplyr::select(S,famsup,Medu,Fedu,guardian)
#Then split the data into two groups, one as training group the other is the test group
## 75% of the sample size
smp_size <- floor(0.7 * nrow(FE))
## set the seed to make your partition reproductible
set.seed(123)
train_ind <- sample(seq_len(nrow(FE)), size = smp_size)
FE.TR <- FE[train_ind, ]
FE.TE <- FE[-train_ind, ]

#now I'm try to build a classification tree
c.tree1 <-  rpart(famsup ~ Medu + Fedu + guardian, method="class", data=FE.TR ,control=rpart.control(minsplit = 1,minbucket = 1,cp=0.001))
printcp(c.tree1)
#CP nsplit rel error  xerror     xstd
#1 0.0216216      0   1.00000 1.00000 0.049320
#2 0.0115830      5   0.89189 0.91120 0.048291
#3 0.0038610      6   0.88031 0.91506 0.048341
#4 0.0015444     10   0.86486 0.94595 0.048724
#5 0.0012870     18   0.84942 0.96911 0.048990
#6 0.0010000     21   0.84556 0.96911 0.048990

#the error is a little too high I tried to set a lower cp
c.tree2 <-  rpart(famsup ~ Medu + Fedu + guardian, method="class", data=FE.TR,control=rpart.control(minsplit = 1,minbucket = 1,cp=0.000001))
printcp(c.tree2)
#          CP nsplit rel error  xerror     xstd
#1 0.02162162      0   1.00000 1.00000 0.049320
#2 0.01158301      5   0.89189 0.94208 0.048678
#3 0.00386100      6   0.88031 0.89575 0.048086
#4 0.00154440     10   0.86486 0.91892 0.048391
#5 0.00128700     18   0.84942 0.96911 0.048990
#6 0.00096525     21   0.84556 0.96911 0.048990
#7 0.00000100     25   0.84170 0.96525 0.048947

# the error is lower but sacrifising part of the complexity cost.

post(c.tree2, file = "tree2.ps", title = "education support from family: 1 -  yes, 2 - no")

```
I will now use this classification tree to test the test group
```{r}
FE.TE$predict <- predict(c.tree2, FE.TE, type = "class")
table(FE.TE$famsup, FE.TE$predict)
#      no yes
# no   28  90
# yes  22 160
error=(22+90)/(28+90+22+160)
#0.3733333
```
Turns out the error is pretty low. Thus, the family support could be predicted by the parents' education together with who is the gaurdian

